Non-word repetition in Yélî Dnye

### 2020-09-07 manuscript.Rmd
add minor M2A comments from reading

@@

Naturally, designing the task in this way may render the study of individual variaiton within the population  more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit.

@@

- Does the frequency of sounds across languages predict NWR? Are rarer sounds more often substituted by commoner sounds?
- How does score change as a function of item length in number of syllables?
- Is individual variation attributable to child age, sex, birth order, monolingual status, and/or parental education?

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether.

@@

A script was written to randomize all tokens from all children, pairing each with the auditory target the child had been provided with.  A native research assistant then listened through all productions of a given target (randomized across children and repetition order), and made a judgment of whether the item was correctly or incorrectly repeated. She additionally transcribed exactly what the child said, providing some examples of the types of errors children in general make (without making specific reference to Yélî sounds or the items in the elicitation sets).

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level accuracy [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those.

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number. * **mc** do you agree with this decision or shall we do the alternative?*

>>>>

<!--M2A: I feel like there's a bit of a jump here from explaining what NWR is and what it's used for to saying we used it in the current study to test three things. We should probably have a paragraph justifying its use in this context, no?-->

@@

Naturally, designing the task in this way may render the study of individual variaiton within the population  more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit. <!--M2self: make sure it's apparent how/why a visit is limited-->

@@

- Does the frequency of sounds across languages predict NWR? Are rarer sounds more often substituted by commoner sounds?
- How does score change as a function of item length in number of syllables?
- Is individual variation attributable to child age, sex, birth order, monolingual status, and/or parental education?
<!--M2A: I'm imagining a reviewer who's knowlegeable about phonology and child development but who hasn't encountered NWR before. I think it would be beneficial to provide a concrete example of a nonword and an inexact repetition above so that they can conceive of what these questions imply (analytically) before we get into the methodological weeds.-->

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.
 <!--M2A: I didn't look for the SM yet, but probably best not to provide exact hamlet names-->

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether. 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->

@@

A script was written to randomize all tokens from all children, pairing each with the auditory target the child had been provided with.  A native research assistant then listened through all productions of a given target (randomized across children and repetition order), and made a judgment of whether the item was correctly or incorrectly repeated. She additionally orthographically transcribed exactly what the child said, providing some examples of the types of errors children in general make (without making specific reference to Yélî sounds or the items in the elicitation sets).

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level accuracy [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those.
<!--M2A: A concrete scoring example would be a good idea here-->

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.
<!--
A2M: do you agree with this decision or shall we do the alternative?
M2A: yeah, i think adding abirtrarily small frequencies to the NA cases is fraught for the reasons we talked about this morning (i.e., possible systematic undersampling in phoible that can't be helped).
-->
